\section{Weighted Fair Queueing}

\subsection{Max-Min Fairness}

\begin{itemize}
	\item Purpose - Allocate fair share of resource among competing req.
	\item Suppose:
	\begin{itemize}
		\item R units of resource
		\item n requests
		\item Fair allocation R/n?
		\item Some requests for less?
		\item How to apportion excess?
	\end{itemize}
	\item Assume n requests for $d_1, d_2, \ldots d_n$, with $D =
		\sum_{i=1}^{n} d_i$ and $d_j < d_j+1$
	\item Assume $D>R$
	 \begin{itemize}
		 \item $r_1 = min(D_1,\frac{R}{n})$
	\end{itemize}
	\item Now remaining level of resource is $R-r_1$, allocate:
	\begin{itemize}
		\item $r_2 = min(d_2, \frac{R-r_1}{n-1})$ etc.
	\end{itemize}
	\item Two important Features:
	\begin{itemize}
		\item Max num of req. met in full, starting with min request
		\item Users which req more of resource can be allocaed by alg to
			receive same allocation
	\end{itemize}
	\item Equivalent Algorithm:
	\begin{itemize}
		\item Divide R evenly among req.
		\item For those req. which have received more than required:
		\begin{itemize}
			\item Divide excess evenly among outstanding req.
		\end{itemize}
		\item Repeat until no excess resources to redistribute
	\end{itemize}
\end{itemize}

\subsection{Quality of Service Guarantees}

\begin{itemize}
	\item FCFS doesn't guarantee BW share to flow
	\item Use Per-Flow queueing
	\item Principle - Generalised Processor Sharing
	\begin{itemize}
		\item Assume traffic on flow is a fluid, can be continuously
			subdivided, is possible to offer service to
			infinitesimal quantity of traffic on a flow
		\item Can devise a scheme to give exactly req. proportion of BW
			to each flow
		\item Practical systems, pkts indivisible, can only approximate
	\end{itemize}
\end{itemize}

\subsection{Generalised Processor Sharing}

\begin{itemize}
	\item Aka Fluid-Flow Fair Queuing
	\item Offers max-min fiar share of BW to each flow
	\item When all flows busy, each gets req. share of avail. BW
	\item When only some flows busy BW share of idle flows allocated in
		max-min fair way to backlogged flows
	\item Let $S_i(t, t +\delta t)$ be amnt data from flow i served in
		interval $[t, t+\delta t]$
	\item If flow i backlogged during interval, for any flow j:
	\begin{itemize}
		\item $\frac{S_i(t, t+\delta t)}{S_j(t, t+\delta t)} \ge
			\frac{\delta_i}{\delta_j}$
	\end{itemize}
\end{itemize}

\subsection{Weigthed Fair Queuing}

\begin{itemize}
	\item Round Time units - bits served per flow
	\item As no. backlogged flows changes, actual time taken to serce bit
		changes in inverse proportion
	\item So far giving each flow equal BW share
	\item Let $F_{i,k}$ and  ${L_i,k}$ be the finish/length resp. of the
		k-th pkt to arrive on flow i
	\item Let $R_{i,k}$ be the round in which it arrived. Evidently:
	\begin{itemize}
		 \item $F_{i,k}=max(F_{i,k-1},R_{i,k})+L_{i,k}$
	\end{itemize}
	\item Problem: calculation of $R_{i,k}$
	\item Need to know relationshiop between R(t) and t when pkt arrives
	\item In principle:
	\begin{itemize}
		\item $R(t_\delta t)=R_(t)+\frac{\delta t}{B_(t)}$
		\item if B(t) is constant in interval $[t,t+ \delta t]$
		\item Need to keep track of when B(t) changes
		\begin{itemize}
			\item When pkts arrive/depart
		\end{itemize}
	\end{itemize}
	\item Let E(t) be time of last ecent before time t (arrival/depart)
	\item R(E(t)) is round no. when last event occurred.
	\item Suppose that at time t, pkt arrives, then:
	\begin{itemize}
		\item $R(t) <- R(E(t)) + \frac{t-E(t)}{B(t)}$ //Update value of
			R(t)
		\item If pkt arrives into empty buffer:
		\item $B(t) <- B(E(t))+1$ // Now 1 more backlogged flow
		\item Else
		\item $B(t) <- B(E(t))
		\item end if
		\item E(t) <- t$
	\end{itemize}
	\item Suppose that at time t, pkt departs GPS discipline then:
	\begin{itemize}
		\item $R(t) <- R(E(t))+\frac{t-E(t)}{B(t)}$
		\item If pkt leaves behind empty buffer then:
		\item $B(t) <- B(E(t)) - 1$
		\item else
		\item B(t) <- B(E(t))
		\item end if
		\item E(t) <- t
	\end{itemize}
	\item If all calc done when pkt arrives, alg is:
	\begin{enumerate}
		\item Calc tentative value for R(t) (assuming no departures
			since E(t)
		\item Find pkt with min finish time $F_{min}$ and check if
			$F_{min}< R(t)$
	\end{enumerate}
	\item If so:
	\begin{itemize}
		\item Calc its time of departure t as
			$E(t)+B(E(t))(F_{min}-R(E(t)))$
		\item Process departure as described earlier ($R(t)=F_{min}$)
		\item $E(t)<-t$
		 \item Go back to step 1
	\end{itemize}
	\item Otherwise:
	\begin{itemize}
		\item Process newly arrived cell as described earlier.
	\end{itemize}
\end{itemize}

\subsection{Weighting the Queues}

\begin{itemize}
	\item Redefine B(t) as:
	\begin{itemize}
		\item $B(t)=\sum_{all Flows (i),active At Time (t)}^{} N\phi_i$
	\end{itemize}
	\item Service time in rounds of pkt is now $\frac{L_{i,k}}{N\phi_i}$
	\item Finish time calc as:
	\begin{itemize}
		\item $F_{i,k}=max(F_{i,k-1},R_{i,k})+\frac{L_{i,k}}{N\phi_i}$
	\end{itemize}
	\item Since N acting as scaling factor, can set to unity:
	\begin{itemize}
		\item $B(t) = \sum_{all Flows (i), active At Time t}^{} \phi_i$
		\item $F_{i,k}=max(F_{i,k-1},R_{i,k}) + \frac{L_{i,k}}{\phi_i}$
	\end{itemize}
\end{itemize}
